---
title: "gradphys"
output: 
  html_document: 
    toc: true
    keep_md: true
    code_folding: hide
    highlight: zenburn
    theme: spacelab
    number_sections: true
date: "2023-12-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(fontawesome)
```

# Background [WIP]

::: {#WIP .alert .alert-warning}
::: {.h3 .alert-heading .alert-warning}
`r fontawesome::fa('triangle-exclamation')` **Work in Progress**
:::

------------------------------------------------------------------------

**This article is currently a work in progress. Feel free to observe while it gets filled out!**
:::

# Data

## Introduction

Data were gathered for the years **2015**-**2022**, accessed on 21 Dec 2023, in Excel spreadsheets available from the
***American Institute of Physics*** (**AIP**); via their [Roster of Physics Departments with Enrollment and Degree
Data](https://www.aip.org/taxonomy/term/220). Due to the structure of the data reporting, enrollment data are considered
aggregated for the traditional academic year (beginning in the Fall of the previous year and ending in the Spring of the
listed year); in addition to an aggregate of all degree recipients, awarded or extant, for the calendar extent (Jan ->
Dec) of the listed year.

For example, in the 2015 report, data regarding enrollments are the finalized numbers of the 2014-2015 academic year;
plus all conferred Bachelors, MS, and PhD certificates from January through December (inclusive) of the 2015 calendar
year.

Datasheets were edited at an intermediary step to unify and homogenize data condensation into a
[`tidy`](https://tidyr.tidyverse.org/) data set. Prior to **2017**, the `Highest Physics Degree Offered` field and some
associated enrollment data was not available in the survey. In our analysis, data were infilled on the condition that if
a `Masters` or `PhD` degree certificate was conferred during those years, then the corresponding program at the
appropriate level must have existed. The `Notes` annotations field was removed from this analysis, as the field appeared
consistent across years, and after 2019 split into a separate datasheet of the same annual report- the `Notes` values
and changelog are available at the source data. `Institution`-level data were adjusted for varying
spellings/canonization over time (e.g. `Appl Phy` -\> `Appl Phys`, `Coll of -`, etc.).  Where possible, an
`Institution`s name as of 2022 is used. 

and (`TODO :>`) `Coll -> U` if the institution re-designated in a later year.

#### TODO :>
Known issues: 
- Multiple institutions failed to report in 2016/2020. 
- Mary Baldwin Coll -> U in 2017
- Piedmont Coll -> U in 2022
- William & Mary, namin Coll of -
- Purdue U-West Lafayette should be marked as Combined astro program
- SUNY Brockport

The process used to incorporate this data is encoded in a function, `process_data(...)`, demonstrated in the following R
script:

```{r, file=c('process_data.R')}
```

## Table

Columns have summary metrics generated for groupings of observations, given an `Institution` over a set of `Year`s.
These have been marked as **Accumulative** or **Averaged**. Data at the individual observation level (foreach
`Institution`, foreach `Year`) ***have not yet been altered*** from their original values.

- TODO :> make all-NA years orange/red bckgrnd

```{r message=F, warning=F, message=F, paged.print=TRUE}
library(tidyverse)
library(reactable)
library(ggplot2)
library(plotly)
library(htmltools)
library(crosstalk)
DATA_DIR = './data/xlsx_edit/'
source('./process_data.R', local = knitr::knit_global())
##########
##-----##
##----##
##---##   Data Table
##----##
##-----##
##########
rosterdata <- process_data(DATA_DIR);
tdata <- SharedData$new(rosterdata);

#####
##  Generate Reactable data table
#####
rxtbl <- tdata |>
    reactable(
        ####
        ##  main table
        ####
        elementId = 'rosterphys14_22-tbl',
        filterable = T,
        #searchable = T,
        groupBy = c('Institution'),
        bordered = T,
        striped = T,
        highlight = T,
        compact = T,
        pagination = F,
        showPageSizeOptions = T,
        pageSizeOptions = c(6, 12, 18),
        paginationType = "jump",
        height = 768,
        defaultColDef = colDef(
            header = function(value) gsub(".", " ", value, fixed = TRUE),
            headerClass = "sticky",
            cell = function(value) format(value, nsmall = 1),
            align = "center",
            minWidth = 100,
            headerStyle = list(background = "#f7f7f8"),
            vAlign = 'center',
            headerVAlign = 'bottom',
            format = colFormat(
                separators = T,
                digits = 0
            )
        ),
        defaultSorted = list(
            Year = 'asc',
            Physics_PhDs = 'desc',
            Physics_Bachelors = 'desc'
        ),
        defaultPageSize = 6,
        minRows = 4,
        ####
        ##  Per-Column defns
        ####
        columns = list(
          `Year` = colDef(
            name = 'Year',
            align = "center",
            minWidth = 60,
            sticky = "left",
            sortable = T,
            defaultSortOrder = "asc"
            ),
          `Institution` = colDef(
            name = 'Institution',
            align = 'left',
            minWidth = 240,
            sticky = "left"
            ),
          `State` = colDef(
            name = 'State',
            align = 'center',
            ),
          `Highest_Physics_Degree_Offered` = colDef(
            name = "Highest Physics Degree Offered",
            aggregate = "unique",
            align = 'center',
            ),
          `Fall_Total_Graduate_Student_Enrollments` = colDef(
            name = "Total Grad Student Enrollment (Fall)\n(Averaged)",
            aggregate = "mean",
            align = 'left',
            ),
          `Physics_PhDs` = colDef(
            name = "Physics PhDs\n(Accum)",
            aggregate = "sum",
            align = 'center',
            format = colFormat(digits=0)
            ),
          `Exiting_Physics_Masters` = colDef(
            name = "Exiting Physics Masters\n(Accum)",
            aggregate = "sum",
            align = 'center',
            format = colFormat(digits=0)
            ),
          `Fall_FirstYear_Graduate_Student_Enrollments` = colDef(
            name = "First-Year Grad Student Enrollment (Fall)\n(Averaged)",
            aggregate = "mean",
            align = "left",
            ),
          `Physics_Bachelors` = colDef(
            name = "Physics Bachelors\n(Accum)",
            aggregate = "sum",
            align = "center",
            format = colFormat(digits=0)
            ),
          `Fall_Senior_Enrollments` = colDef(
            name = "Senior Enrollment (Fall)\n(Averaged)",
            aggregate = "mean",
            align = "center",
            ),
          `Fall_Junior_Enrollments` = colDef(
            name = "Junior Enrollment (Fall)\n(Averaged)",
            aggregate = "mean",
            align = "center",
            ),
          `FirstTerm_Introductory_Physics_Course_Enrollments` = colDef(
            name = "First-Term Intro Physics Course Enrollment\n(Averaged)",
            aggregate = "mean",
            align = "left"
            ),
          `FirstTerm_Introductory_Physical_Science_and_Astronomy_Course_Enrollments` = colDef(
            name = "First-Term Intro Phys. Sci. and Astro. Course Enrollment\n(Averaged)",
            aggregate = "mean",
            align = "left"
            ),
          `Fall_NonUS_Graduate_Student_Enrollments` = colDef(
            name = "Non-US Grad Student Enrollment (Fall)\n(Averaged)",
            aggregate = "mean",
            align = "left"
            ),
          `Astro_Program` = colDef(
            name = "Astro Program?",
            align = "right",
            aggregate = "unique"
            )
        )
    )

#####
##  Display Shiny with Crosstalk/htmltools widgets
#####
shiny::fluidRow(
  shiny::column(
    2,
    htmltools::browsable(
      tagList(
        tags$div(
          class = 'd-grid gap-2 mx-auto',
          # CSV download button
          tags$button(
            tagList(fontawesome::fa("download"), "\tCSV Data"),
            class = 'btn btn-success',
            onclick = "Reactable.downloadDataCSV('rosterphys14_22-tbl', 'rosterphys14_22.csv')"
          ),
          # Expand/Collapse button
          tags$button(
            "Expand/Collapse",
            class = 'btn btn-info',
            onclick = "Reactable.toggleAllRowsExpanded('rosterphys14_22-tbl')",
          ),
        )
      )
    ),
    # filter-by degree
    filter_checkbox("degree", "Degree Level", tdata, ~Highest_Physics_Degree_Offered),
    # TODO :> allow column selection in this menu
    # Group Selection
    htmltools::browsable(
        tagList(
            # Group-by selection
            div(tags$label("Group by", `for` = "rosterphys14_22-select")),
            tags$select(
              id = "rosterphys14_22-select",
              onchange = "Reactable.setGroupBy('rosterphys14_22-tbl', this.value ? [this.value] : [])",
              tags$option("None", value = ""),
              tags$option("Institution", value = "Institution"),
              tags$option("Program", value = "Highest_Physics_Degree_Offered"),
              tags$option("State", value = "State"),
            ),
        )
    ),
  ),
  shiny::column(
    10,
    shiny::fluidRow(
      htmltools::browsable(
        tagList(
          tags$div(
          )
        )
      ),
    ),
    # DataTable
    rxtbl
  )
)
```

## Observations

Multiple notes arise from a brief overview of this collection:

### Summary of Degree Type by Unique Institutions

```{r message=FALSE, warning=FALSE, paged.print=TRUE}
print('TODO:> .')
```

- A natural boundary to delineate the `r length(unique(rosterdata$Institution))` unique `Institutions` in this set is in
  the categorical division of `Highest Physics Degree Offered` (`BS`, `MS`, and `PhD`) in the most recent year of the
  roster publication. We will return to this Bayesian decision boundary of the search space later. Furthermore, we will
  consider `PhD`-granting `Institution`s as separate from the combined group of `BS/MS`-granting `Institution`s

### PhD Availability by Region

```{r message=FALSE, warning=FALSE, paged.print=TRUE}
print('TODO:> .')
```

- There is at least one Institution in every state and territory, except the US Virgin Islands, that has a Physics PhD
  program. There are `TODO` unique `Institutions` offering a `PhD` as of `2022`.
  
- `TODO :>` Use shapefiles to demonstrate gradient of universities and availability of programs
  - Maybe leave this summary graphic until County-level data is established.

### Temporal Changes of Observations

```{r message=FALSE, warning=FALSE, paged.print=TRUE}
print('TODO:> .')
```

- The University of Minnesota (MN) - Minneapolis offered a PhD through **2015-2019** (the past extent of the dataset), but
  is no longer part of the reported data set. However, the University of Minnesota - Twin Cities in **2020-Current**,
  despite having no data prior to 2020, produced similar numbers to the prior Institution. These two should be merged for
  the purposes of this analysis, as if the Physics department is continuous through these years across both Institutions.
  Similar behavior occurs in a handful of other institutions.

### Applied Physics Departments

```{r message=FALSE, warning=FALSE, paged.print=TRUE}
  print('TODO:> .')
```

- Applied Physics PhDs are granted by a separate department housed in another College, of Engineering or similar,
  complementary to an Institution's existing Physics department. Only `6` Institutions of this type (for our analysis)
  exist: specifically `Stanford`, `Yale`, `Univ. Of Michigan - Ann Arbor`, `Columbia`, `Cornell`, and `Rice` 
  Universities.

### Top 10 States by accumulative `Physics PhDs`

```{r message=FALSE, warning=FALSE, paged.print=TRUE}
print('TODO:> .')
```

- The Top 10 ranking of cumulative production since 2015 of Physics PhDs is, in descending order: `CA` (1926), `NY`
  (1154), `MA` (1105), `TX` (907), `IL` (700), `FL` (700), `PA` (623), `OH` (599), `MI` (530), and `CO` (419). We should
  ensure we perform a population-corrected analysis of the per-capita rate of completing a PhD in these states.
  Interestingly, this ranking is disjoint from the same ordering for the production of Physics Bachelors. Yet another
  ranking is produced when ordered by descending quantity of the three degree levels (`BS`, `MS`, and `PhD`).

### Top 20 Institutions by accumulative `Physics_PhDs`

```{r message=FALSE, warning=FALSE, paged.print=TRUE}
# TODO :> TOP 20 ranking
```

-   The Top 20 ranking Institutions of cumulative production over the past decade ordered by `PhD`, alongside 
    accumulative `BS` degrees conferred are:

| **`Institution`**           | **`Physics PhDs`** | `Physics Bachelors` |
|-----------------------------|:------------------:|:-------------------:|
| `CA-U of, Berkeley`         |        286         |         931         |
| `IL-U of, Urbana/Champaign` |        270         |        1129         |
| `Colorado-U of, Boulder`    |        263         |         663         |
| `Harvard U`                 |        262         |         423         |
| `Ohio State U`              |        252         |         586         |
| `Maryland-U of, Coll Park`  |        241         |         576         |
| `Stanford U`                |        225         |         197         |
| `Texas-U of, at Austin`     |        186         |         635         |
| `Mass Inst of Tech (MIT)`   |        184         |         376         |
| `Cornell U`                 |        182         |         368         |
| `Chicago-U of`              |        181         |         449         |
| `WI-U of, Madison`          |        181         |         387         |
| `Michigan-U of, Ann Arbor`  |        180         |         439         |
| `CA-U of, Los Angeles`      |        170         |         644         |
| `SUNY-Stony Brook`          |        169         |         423         |
| `CA-U of, San Diego`        |        169         |         405         |
| `Texas A&M-College Station` |        168         |         265         |
| `Washington-U of`           |        163         |        1286         |
| `Princeton U`               |        161         |         203         |
| `Florida State U`           |        160         |         165         |

-   `TODO :>` How many departments inconsistently reported data?

-   `TODO :>` How many departments ceased reporting per each year, any departments end their `PhD` or `BS` program?

-   `TODO :>` COVID-19 impacts?

-   `TODO :>` Frequency Distribution of PhDs

-   `TODO :>` Top 10 by accumulative Physics Bachelors?
    Frequency Distribution?

    -   what the f is going on with the higher number of exiting physics masters (like U Washington)?

-   `TODO :>` Filter by latest year reported + frequency graph/histogram by year, earliest year reported, all NA (e.g., 
    dead departments, no reports, new depts).

-   `TODO :>` What is the boundary/difference in Phys bachelors production and undergrad enrollment figures between 
    programs that offer BS/MS as highest vs PhD as highest?

## Targeted Adjustments

-   flow across all years to fix misspells & names

## Additional Data

-   Faculty count?

-   Institution-level COMMON dataset

-   Voronoi-Cell approximation of nearest PhD or BS

-   Department Specialties? <- n-gram analysis

-   Undergraduate population total enrollments?

-   County of State that Institution is located in

-   Median Salary of surrounding

-   Percentile of Median Salary of county relative to State

-   Percentile median salary of county Relative to Country?

-   Gini Index.

-   Median Rent of surrounding

-   \% vote in last election?

-   Political Party of State governor?

-   Online Program availability?

-   Has nearby or attached National Lab (`as.factor(...)` with Y/N levels; sep col of assoc. lab)

-   Link to faculty/dept page

-   FAFSA ID

-   R1 research designation

------------------------------------------------------------------------

# Methods

## Problem-Domain Adjusted and Summary Metrics

-   `TODO :>` Ratio of Accumulative PhDs produced over timeframe to average number of graduate students enrolled per year

-   `TODO :>` Interpolation across NA rows?

### Resiliency and Conversion Scores

-   `TODO :>` Hertzsprung-Russell Diagram here [https://observablehq.com/\@d3/hertzsprung-russell-diagram?intent=fork](https://observablehq.com/@d3/hertzsprung-russell-diagram?intent=fork){.uri}

### Per-Capita PhDs Conferred vs. Undergrad Physics Enrollment Population

### Altered Data

------------------------------------------------------------------------

## Time-Series Regression

## `XGBoost` Random Forest

------------------------------------------------------------------------

## Bollinger-Bands Volatility Metric

------------------------------------------------------------------------

# Analysis

------------------------------------------------------------------------

# Results

------------------------------------------------------------------------

# Future Work

------------------------------------------------------------------------

# References
